Supplemental Material: Multidimensional analysis and detection of informative features in human brain white matter

### 2021-04-20 reports/manuscript/paper.tex

Data from the ALS and WH studies was processed in a similar manner,
using the Matlab Automated Fiber Quantification toolbox (\texttt{mAFQ})
\cite{yeatman2012tract}: streamlines representing fascicles of white
matter tracts were generated using a determinstic tractography algorithm
that follows the prinicpal diffusion direction of the diffusion tensor
in each voxel (STT) \cite{basser2000vivo}. Eighteen major
tracts, which are enumerated in the supplemental material, were identified
using multiple criteria: streamlines are selected as candidates for
inclusion in a bundle of streamlines that represents a tract if they
pass through known inclusion ROIs and do not pass through exclusion
ROIs \cite{wakana2007reproducibility}. In addition, a probabilistic
atlas is used to exclude streamlines which are unlikely to be part of
a tract \cite{Hua2008-sh}. Each streamline is resampled to 100 nodes
and the robust mean at each location is calculated by estimating the 3D
covariance of the location of each node and excluding streamlines that
are more than 5 standard deviations from the mean location in any node.
Finally, a bundle profile of tissue properties in each bundle was created
by interpolating the value of MRI maps of these tissue properties to the
location of the nodes of the resampled streamlines designated to each
bundle. In each of 100 nodes, the values are summed across streamlines,
weighting the contribution of each streamline by the inverse of the
mahalanobis distance of the node from the average of
that node across streamlines. This means that streamlines that are more
representative of the tract contribute more to the bundle profile, relative
to streamlines that are on the edge of the tract.

>>>>

Data from the ALS and WH studies was processed in a similar manner,
using the Matlab Automated Fiber Quantification toolbox
(\texttt{mAFQ}, version 1.1 for WH and version 1.2 for ALS)
\cite{yeatman2012tract}: streamlines representing fascicles of white
matter tracts were generated using a determinstic tractography algorithm
that follows the prinicpal diffusion direction of the diffusion tensor
in each voxel (STT) \cite{basser2000vivo}. Eighteen major
tracts, which are enumerated in the supplemental material, were identified
using multiple criteria: streamlines are selected as candidates for
inclusion in a bundle of streamlines that represents a tract if they
pass through known inclusion ROIs and do not pass through exclusion
ROIs \cite{wakana2007reproducibility}. In addition, a probabilistic
atlas is used to exclude streamlines which are unlikely to be part of
a tract \cite{Hua2008-sh}. Each streamline is resampled to 100 nodes
and the robust mean at each location is calculated by estimating the 3D
covariance of the location of each node and excluding streamlines that
are more than 5 standard deviations from the mean location in any node.
Finally, a bundle profile of tissue properties in each bundle was created
by interpolating the value of MRI maps of these tissue properties to the
location of the nodes of the resampled streamlines designated to each
bundle. In each of 100 nodes, the values are summed across streamlines,
weighting the contribution of each streamline by the inverse of the
mahalanobis distance of the node from the average of
that node across streamlines. This means that streamlines that are more
representative of the tract contribute more to the bundle profile, relative
to streamlines that are on the edge of the tract.

### 2021-04-20 reports/manuscript/paper_changes.tex

Data from the ALS and WH studies was processed in a similar manner,
using the Matlab Automated Fiber Quantification toolbox (\texttt{mAFQ})
\cite{yeatman2012tract}: streamlines representing fascicles of white
matter tracts were generated using a determinstic tractography algorithm
that follows the prinicpal diffusion direction of the diffusion tensor
in each voxel (STT) \cite{basser2000vivo}. \replaced{Eighteen major}{Major}
tracts\added{, which are enumerated in the supplemental material,} were identified
using multiple criteria: streamlines are selected as candidates for
inclusion in a bundle of streamlines that represents a tract if they
pass through known inclusion ROIs and do not pass through exclusion
ROIs \cite{wakana2007reproducibility}. In addition, a probabilistic
atlas is used to exclude streamlines which are unlikely to be part of
a tract \cite{Hua2008-sh}. Each streamline is resampled to 100 nodes
and the robust mean at each location is calculated by estimating the 3D
covariance of the location of each node and excluding streamlines that
are more than 5 standard deviations from the mean location in any node.
Finally, a bundle profile of tissue properties in each bundle was created
by interpolating the value of MRI maps of these tissue properties to the
location of the nodes of the resampled streamlines designated to each
bundle. In each of 100 nodes, the values are summed across streamlines,
weighting the contribution of each streamline by the inverse of the
\replaced{mahalanobis}{mahalnobis} distance of the node from the average of
that node across streamlines. This means that streamlines that are more
representative of the tract contribute more to the bundle profile, relative
to streamlines that are on the edge of the tract.

>>>>

Data from the ALS and WH studies was processed in a similar manner,
using the Matlab Automated Fiber Quantification toolbox
(\texttt{mAFQ}\added{, version 1.1 for WH and version 1.2 for ALS})
\cite{yeatman2012tract}: streamlines representing fascicles of white
matter tracts were generated using a determinstic tractography algorithm
that follows the prinicpal diffusion direction of the diffusion tensor
in each voxel (STT) \cite{basser2000vivo}. \replaced{Eighteen major}{Major}
tracts\added{, which are enumerated in the supplemental material,} were identified
using multiple criteria: streamlines are selected as candidates for
inclusion in a bundle of streamlines that represents a tract if they
pass through known inclusion ROIs and do not pass through exclusion
ROIs \cite{wakana2007reproducibility}. In addition, a probabilistic
atlas is used to exclude streamlines which are unlikely to be part of
a tract \cite{Hua2008-sh}. Each streamline is resampled to 100 nodes
and the robust mean at each location is calculated by estimating the 3D
covariance of the location of each node and excluding streamlines that
are more than 5 standard deviations from the mean location in any node.
Finally, a bundle profile of tissue properties in each bundle was created
by interpolating the value of MRI maps of these tissue properties to the
location of the nodes of the resampled streamlines designated to each
bundle. In each of 100 nodes, the values are summed across streamlines,
weighting the contribution of each streamline by the inverse of the
\replaced{mahalanobis}{mahalnobis} distance of the node from the average of
that node across streamlines. This means that streamlines that are more
representative of the tract contribute more to the bundle profile, relative
to streamlines that are on the edge of the tract.
